Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2020-04-28 article.tex
Mention transition counts

TMHs, on the other hand, also have their function constraints, 
yet can occur multiple time a pathogen's proteome.
One can safely assume a pathogen's proteome contains multiple TMHs.
Therefore, it may be beneficial for the host
if its immune system would be more attentive towards TMHs.
And maybe this has already happened: MHC-I already detects hydrophobic
polypeptides. This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focussing on TMHs only, if a signal of selection
can be detected.

@@

When using DNA sequences, one can use dS/dN rates 
to detect the signature of selection \cite{murrell2015gene}.
Using AA sequences, however, has its advantages,
as its is closer to the actual phenotype
selection acts upon: DNA may never be translated to RNA,
or its RNA may never be transcribed \cite{diz2012proteomics}
\richel{improve upon this point}.
When using a preoteome in phylogenetic research,
we know that the majority of proteins are selected to just 
maintain their function most of the time, where
is the timespans there is selection, only a few AAs
can actually increase the 'fitness' of the 
protein \cite{anisimova2009investigating}.
There, when generalizing the dynamics of mostly purifying selection (to maintain
a protein's function) and a short duration of positive selection,
those genes that are selected cannot be detected \cite{yang2000statistical}.

\paragraph{Amino acid substitution models}

One important assumption for such phylogenetic inference
is the amino acid substitution model (AASM), which may be different
in TMHs. An AASM encodes the relative rates at which amino acids
change through evolutionary time. Some amino acids have more 
DNA triplets encoding for them, which is reflected in the AASM.
Already for decades, AASMs are devised and refined from whole proteomes,
such as the Blosum62 (\cite{henikoff1992amino}), 
Dayhoff (\cite{dayhoff197822}), JTT (\cite{jones1992rapid}), 
and WAG (\cite{whelan2001general}) models. 
That such a generalization
is not always desired, AASMs are devised for 
chloroplasts (the CPREV model, see \cite{adachi2000plastid})
and mitochondria (the MTREV24 model, see \cite{adachi1996model}).
In all AASM, TMHs are generalized out and therefore, 
there is no specialized AASM for TMHS. It would, however,
be possible to derive an AASM from in silico predictions. 
An in silico prediction would bypass any effects of evolution.
Another AASM could be constructed from humans, if we assume
that there is no selection on TMHs, except for being functional.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{In silico prediction of TMH AASM}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

We use a Monte-Carlo approach to produce an AASM for TMHs.
We simulate a random amino acid sequence with a length 23.
Using TMHMM, we determine if a sequence encodes for a TMH.
For a TMH sequence, we independently change 
each locus to all other amino acids.
Of each mutation, we count how often the new sequence remains a TMH.
Doing this for many random amino acid sequences, we can determine
the AASM. See figure \ref{fig:tmh_aasm_method} for a schematic
overview of this method.

\begin{figure}[!htbp]
  \includegraphics[width=0.9\textwidth]{tmh_aasm_method.png}
  \caption{
    In silico prediction of TMH AASM
  }
  \label{fig:tmh_aasm_method}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Baseline TMH mutation rate}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

	% \frans{
	%   Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
	%   Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch? 
	% }
	% The evolutionary conservation of TMHs was measured from a
	% DNA alignment of multiple transmembrane proteins in multiple
	% species of the mycobacterium bacterial family
	% \richel{how obtained exactly? how alignment done?}
	% \frans{
	% Via entrez en blast 
	%   Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
	% }
	% .
	% Using the \verb;tmhprot; R package (\cite{tmhprot}), the DNA alignment is split
	% into two alignments, one for the TMH parts, another for the non-TMH
	% parts. Each alignment was tested by the \verb;mcbette; 
	% R package (\cite{mcbette}) to select the Bayesian inference model with
	% the highest evidence (a.k.a. the marginal likelihood) using the nested
	% sampling approach as described in \cite{maturana2018model},
	% using the popular Bayesian phylogenetic tool 
	% BEAST2 (\cite{bouckaert2014beast}) in the back-end.
	% \verb;mcbette; used a set of 40 candidate inference models, 
	% consisting of all combinations of 
	% 4 site models (JC, HKY, TN, GTR), 
	% 2 clock models (strict, RLN) and 
	% 5 tree priors (Yule, BD, CBS, CCP, CEP).

	% For each alignment, the inference model with the highest evidence
	% is used in a Bayesian inference. From each Bayesian posterior,
	% the parameter estimates regarding mutation rates (including clock rate)
	% were obtained and compared to determine which realized mutation rate is lower.
	% The Markov chain Monte Carlo was set up in such a way that the effective sample
	% size for the likelihood of the inference model is above the recommended value
	% of 200 (\cite{bouckaert2014beast}).

@@

@@

>>>>

TMHs, on the other hand, also have their function constraints, 
yet can occur multiple time a pathogen's proteome.
One can safely assume a pathogen's proteome contains multiple TMHs.
Therefore, it may be beneficial for the host
if its immune system would be more attentive towards TMHs.
And maybe this has already happened: MHC-I already detects hydrophobic
polypeptides. This feature, however, may also be caused by selection
to detect hydrophobic regions in the soluble proteins of pathogens.
It is unknown, when focusing on TMHs only, if a signal of selection
can be detected.

@@

When using DNA sequences, one can use $dS/dN$ rates 
to detect the signature of selection \cite{murrell2015gene}.
Using AA sequences, however, has its advantages,
as its is closer to the actual phenotype
selection acts upon: DNA may never be translated to RNA,
or its RNA may never be transcribed \cite{diz2012proteomics}
\richel{improve upon this point}.
When using a proteome in phylogenetic research,
we know that the majority of proteins are selected to just 
maintain their function most of the time, where
is the time spans there is selection, only a few AAs
can actually increase the 'fitness' of the 
protein \cite{anisimova2009investigating}.
There, when generalizing the dynamics of mostly purifying selection (to maintain
a protein's function) and a short duration of positive selection,
those genes that are selected cannot be detected \cite{yang2000statistical}.

@@

@@

	% \frans{
	%   Ik heb zitten denken over de evolutie naast conservering is occurrence ook heel belangrijk. Moeten we de mate van conservering niet wegen ten opzichte van de occurrence.
	%   Wat ik daarmee bedoel is data als je kiest voor een mega geconseveerd epitope maar van dat type komen er maar 40 in een proteome voor versus een minder geconserveerd epitope maar daar zijn wel 100.000 varianten van dan heb je meer kans om 1 van die 100.000 te pakken toch? 
	% }
	% The evolutionary conservation of TMHs was measured from a
	% DNA alignment of multiple transmembrane proteins in multiple
	% species of the mycobacterium bacterial family
	% \richel{how obtained exactly? how alignment done?}
	% \frans{
	% Via entrez en blast 
	%   Ik he been transcript van Maxim ik zal dit proberen te vertalen naar een stukje tekst. 
	% }

@@

% Poisson binomial distribution

@@

\begin{figure}[!htbp]
  \includegraphics[width=\textwidth]{transitions.png}
  \caption{
    Transitions in TMHs that occur more often then expected by chance,
    for a confidence interval of 5 percent.
    Table \ref{tab:transitions} shows the transition counts.
    \richel{
      stub figure
    }
  }
  \label{fig:transitions}
\end{figure}

\begin{table}[!htbp]
  \input{table_transitions.latex}
  \caption{
    Transitions counts, where the row indicates the source state,
    and the column indicates the target state.
    First number per cell is the observed number of this state transition,
    where the second number is the expected number of this state transition
    as predicted by chance.
    An asterisk behind the observed count indicates that this count
    is unlikely to be caused by chance only.
    Figure \ref{fig:transitions} shows which transition counts are 
    unlikely to be caused by chance only.
  }
  \label{tab:transitions}
\end{table}
